% First parameter can be changed eg to "Glossary" or something.
% Second parameter is the max length of bold terms.
\begin{mclistof}{List of Abbreviations}{3.2cm}

\item[CGAP] Consultative Group to Assist the Poor.

\item[DEA] Data Envelopment Analysis.

\item[DMFI] Deposit Taking Microfinance Institution.

\item[EAP] East Asia and the Pacific.

\item[ECA] East and Central Asia.

\item[FinTech] Financial Technology.

\item[FSB] Financial Stability Board.

\item[FSPs] Financial Services Providers.

\item[GNI] Gross National Income.

\item[IFC] International Finance Corporation.

\item[IMF] International Monetary Fund.

\item[IPOs] Initial Public Offers.

\item[KIVA] KIVA Microfund.

\item[KREP] Kenya Rural Enterprise program.

\item[LAC] Latin America and the Caribbean.

\item[MENA] Middle East and North Africa.

\item[MF] Microfinance.

\item[MFI(s)] Microfinance Institution(s).

\item[MIS] Management Information Systems.

\item[MIX] Microfinance Information Exchange.

\item[NBFIs] Non-Bank Financial Institutions.

\item[NDMFI] Non-Deposit Taking Microfinance Institution.

\item[NGO] Non-Governmental Organisation.

\item[OIBM] Opportunity International Bank of Malawi.

\item[OLS] Ordinary Least Squares.

\item[OL-SASL] Opportunity International Savings and Loans LTD.

\item[OSS] Operational Self-Sufficiency.

\item[RCT] Randomized Control Trials.

\item[RFI] Regulated Financial Institution.

\item[SAPs] Structural Adjustment Programs.

\item[SET] Social Exchange Theory.

\item[SFA] Stochastic Frontier Analysis.

\item[SHGs] Self-Help Groups.

\item[SMEs] Small and Medium Enterprises.

\item[SSA] Sub-Saharan Africa.

\item[USAID] United States Agency for International Development.

\end{mclistof} 
